---
title: "miR-601 and miR-630 targets analysis"
output:
  html_document:
    df_print: paged
---

```{r}
require(tidyverse)
require(org.Hs.eg.db)
require(clusterProfiler)
```

Download gene target tables for miR-601.
```{r}
setwd('C:/Users/patipark/Dropbox/STAR/miRNA')

T601 <- as.data.frame(read.delim('miR-601_targets.txt'))

# change name of column 11 and 12
colnames(T601)[c(11,12)] <- c('CWCS','TCS')

```

```{r}
# remove CWCS = 0
T601$CWCS <- as.numeric(as.character(T601$CWCS))

T601 <- T601 %>% subset(CWCS < 0)

```


```{r}
# convert gene symboles to entrez ids
gene.T601 <- as.character(T601$Target.gene)
gene.T601.df <- clusterProfiler::bitr(gene.T601,
                                 fromType = 'SYMBOL',
                                 toType = c('ENTREZID'),
                                 OrgDb = org.Hs.eg.db)

```

```{r}
go.BP.T601 <- enrichGO(gene          = gene.T601.df$ENTREZID, # always use ENTREZID as gene inputs
                       # universe      = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       ont           = "BP", # BP, MF, CC
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
```

```{r}
go.MF.T601 <- enrichGO(gene          = gene.T601.df$ENTREZID, # always use ENTREZID as gene inputs
                       # universe      = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       ont           = "MF", # BP, MF, CC
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
```

```{r}
go.CC.T601 <- enrichGO(gene          = gene.T601.df$ENTREZID, # always use ENTREZID as gene inputs
                       # universe      = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       ont           = "CC", # BP, MF, CC
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
```

```{r}
# add annotation
go.BP.T601@result$Class <- 'BP'
go.CC.T601@result$Class <- 'CC'
go.MF.T601@result$Class <- 'MF'

# combine table
go.T601 <- rbind(go.BP.T601@result,go.CC.T601@result,go.MF.T601@result)
# filter p.adjust < 0.05
go.T601.pval <- go.T601 %>% 
  subset(p.adjust < 0.05) %>%
  arrange(., p.adjust) %>%
  slice_head(., n = 10)

go.T601.pval$p.adjust <- -log10(go.T601.pval$p.adjust)

go.T601.pval$Term <- paste0('[',go.T601.pval$Class,']',go.T601.pval$Description)

go.T601.pval$Term <- factor(go.T601.pval$Term, levels = rev(go.T601.pval$Term))


```

```{r}

GOTerm.miR601 <- ggplot(go.T601.pval,
       aes(x = p.adjust,
           y = Term)) +
  geom_col() +
  theme_bw() +
  labs(x = expression(-log[10]~p-value),
       y = 'GO Terms',
       title = 'miR-601 targetted genes')

GOTerm.miR601

```

Download gene target tables for miR-630.
```{r}
setwd('C:/Users/patipark/Dropbox/STAR/miRNA')

T630 <- as.data.frame(read.delim('miR-630_targets.txt'))

# change name of column 11 and 12
colnames(T630)[c(11,12)] <- c('CWCS','TCS')

```

```{r}
# remove CWCS = 0
T630$CWCS <- as.numeric(as.character(T630$CWCS))

T630 <- T630 %>% subset(CWCS < 0)

```

```{r}
gene.T630 <- as.character(T630$Target.gene)
gene.T630.df <- clusterProfiler::bitr(gene.T630,
                                 fromType = 'SYMBOL',
                                 toType = c('ENTREZID'),
                                 OrgDb = org.Hs.eg.db)
```

```{r}
go.BP.T630 <- enrichGO(gene          = gene.T630.df$ENTREZID, # always use ENTREZID as gene inputs
                       # universe      = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       ont           = "BP", # BP, MF, CC
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
```

```{r}
go.MF.T630 <- enrichGO(gene          = gene.T630.df$ENTREZID, # always use ENTREZID as gene inputs
                       # universe      = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       ont           = "MF", # BP, MF, CC
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
```

```{r}
go.CC.T630 <- enrichGO(gene          = gene.T630.df$ENTREZID, # always use ENTREZID as gene inputs
                       # universe      = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       ont           = "CC", # BP, MF, CC
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
```

```{r}
# add annotation
go.BP.T630@result$Class <- 'BP'
go.CC.T630@result$Class <- 'CC'
go.MF.T630@result$Class <- 'MF'

# combine table
go.T630 <- rbind(go.BP.T630@result,go.CC.T630@result,go.MF.T630@result)
# filter p.adjust < 0.05
go.T630.pval <- go.T630 %>% 
  subset(p.adjust < 0.05) %>%
  arrange(., p.adjust) %>%
  slice_head(., n = 10)

go.T630.pval$p.adjust <- -log10(go.T630.pval$p.adjust)

go.T630.pval$Term <- paste0('[',go.T630.pval$Class,']',go.T630.pval$Description)

go.T630.pval$Term <- factor(go.T630.pval$Term, levels = rev(go.T630.pval$Term))


```

```{r}
GOTerm.miR630 <- ggplot(go.T630.pval,
       aes(x = p.adjust,
           y = Term)) +
  geom_col() +
  theme_bw() +
  labs(x = expression(-log[10]~p-value),
       y = 'GO Terms',
       title = 'miR-630 targetted genes')

GOTerm.miR630

```

```{r}
require(ggpubr)

Figure7.GO <- ggarrange(GOTerm.miR601,GOTerm.miR630,
          labels = c('A','B'),
          nrow = 2,
          ncol = 1, align = 'v')

Figure7.GO
```

Save GO term outputs.
```{r}
# setwd('C:/Users/patipark/Dropbox/STAR/miRNA')
# 
# write.csv(go.T601,'T601_GOTerm_table.csv',row.names = FALSE)
# 
# write.csv(go.T630,'T603_GOTerm_table.csv',row.names = FALSE)
```


Perform KEGG pathways analysis.
```{r}
kg.T601 <- enrichKEGG(gene         = gene.T601.df$ENTREZID,
                      organism     = 'hsa',
                      pvalueCutoff = 0.01)

# combine table
kg.T601 <- kg.T601@result
# filter p.adjust < 0.05
kg.T601.pval <- kg.T601 %>% 
  # subset(p.adjust < 0.05) %>%
  arrange(., p.adjust) %>%
  slice_head(., n = 10)

kg.T601.pval$p.adjust <- -log10(kg.T601.pval$p.adjust)

kg.T601.pval$Description <- factor(kg.T601.pval$Description, levels = rev(kg.T601.pval$Description))


KEGG.miR601 <- ggplot(kg.T601.pval,
       aes(x = p.adjust,
           y = Description)) +
  geom_col() +
  theme_bw() +
  labs(x = expression(-log[10]~p-value),
       y = 'KEGG Pathways',
       title = 'miR-601 targetted genes')
```

```{r}
kg.T630 <- enrichKEGG(gene         = gene.T630.df$ENTREZID,
                      organism     = 'hsa',
                      pvalueCutoff = 0.01)

# combine table
kg.T630 <- kg.T630@result
# filter p.adjust < 0.05
kg.T630.pval <- kg.T630 %>% 
  # subset(p.adjust < 0.05) %>%
  arrange(., p.adjust) %>%
  slice_head(., n = 10)

kg.T630.pval$p.adjust <- -log10(kg.T630.pval$p.adjust)

kg.T630.pval$Description <- factor(kg.T630.pval$Description, levels = rev(kg.T630.pval$Description))


KEGG.miR630 <- ggplot(kg.T630.pval,
       aes(x = p.adjust,
           y = Description)) +
  geom_col() +
  theme_bw() +
  labs(x = expression(-log[10]~p-value),
       y = 'KEGG Pathways',
       title = 'miR-630 targetted genes')
```

```{r}
Figure7.KEGG <- ggarrange(KEGG.miR601,KEGG.miR630,
          labels = c('A','B'),
          nrow = 2,
          ncol = 1, align = 'v')

Figure7.KEGG
```

```{r}
# setwd('C:/Users/patipark/Dropbox/STAR/miRNA')
# ggsave(Figure7.GO,
#        filename = 'Figure7_GOTerms.pdf',
#        height = 15,
#        width = 24,
#        units = 'cm',
#        device = cairo_pdf)
# dev.off()
```

Create a table of genes overlapping between 2 miRNA target data sets for PPI input data.
```{r}
# common targets between miR-601 and miR-630 targets
A <- intersect(unique(gene.T601.df$ENTREZID),unique(gene.T630.df$ENTREZID))
B <- intersect(unique(gene.T630.df$ENTREZID),unique(gene.T601.df$ENTREZID))
C <- A[B %in% A]

gene.df <- clusterProfiler::bitr(C,
                                 fromType = 'ENTREZID',
                                 toType = c('SYMBOL'),
                                 OrgDb = org.Hs.eg.db)

```

We will create an adjacent matrix.
Download table data obtained from STRINGdb
```{r}
library(readr)
setwd('C:/Users/patipark/Dropbox/STAR/miRNA')
my.string <- read_tsv('string_interactions.tsv')
# my.biomarkers <- read.csv('GeneList_04112018.csv', header = TRUE, stringsAsFactors = FALSE)
# df.explevel <- read.csv('Expression_level_ALL_13102018.csv', header = TRUE, stringsAsFactors = FALSE)
```

Get connected nodes
```{r}
library(tidyverse)
## create edges dataframe
my.edges <- my.string %>% 
  dplyr::select(node1,node2,combined_score) %>%
  subset(combined_score >= 0.9) # nodes with combined score >= 0.9 are used to constructed the network

a <- unique(c(my.edges$node1,my.edges$node2))

## create vertices dataframe
my.vertices <- gene.df %>% dplyr::select(SYMBOL) %>% subset(SYMBOL %in% a)

# part of codes is from 
# https://stackoverflow.com/questions/45003500/create-an-adjacency-matrix-or-list-based-on-shared-properties-from-a-data-frame
library(igraph)
# Create graph
g <- graph_from_data_frame(d = my.edges, 
                           vertices = my.vertices,
                           directed = TRUE)
# Create adjacency matrix from graph
M <- as_adjacency_matrix(g, attr = 'combined_score')

# coerces the data set as a matrix
m <- as.matrix(M) 
# this will create an 'igraph object'
gg <- graph.adjacency(m,
                      mode = 'undirected', 
                      weighted = 'combined_score' # use 'combined_socre' as weight
                      )

# change layout algorithm
# my.layout <- layout.reingold.tilford(gg, circular = TRUE)
my.layout <- layout.kamada.kawai(gg)

# convert igraph object to coordinates for building a network using ggplot2
library(intergraph)
g2 <- asNetwork(gg)
library(ggnetwork)
g3 <- ggnetwork(g2,
                layout = my.layout,
                cell.jitter = 0.75)

g3.1 <- g3[!duplicated(g3$vertex.names),]

```

Plot PPI network.
```{r}
library(ggplot2)
ppi <- ggplot(g3,
              aes(x = x, 
                  y = y, 
                  xend = xend, 
                  yend = yend)) +
  geom_edges(aes(linetype = 'solid'),
             #arrow = arrow(length = unit(6, "pt"), type = "closed") #if directed
             color = 'grey50',
             #curvature = 0.2,
             alpha = 0.5) +
  geom_nodes(fill = 'gray', 
             shape = 21,stroke = 0.1,
             size = 3) + 
  geom_text(data = g3.1,
                  aes(x = x,
                      y = y,
                      label = vertex.names), 
                  size = 0.75,
                  segment.color = NA) +
  # scale_colour_manual(values = cols, name = 'Expresion level \n(z-score)') +
  # facet_grid(cellsub ~ cell, scales = 'fixed') + #, scales = 'free', space = 'free') +
  coord_fixed(ratio = 1) +
  # theme_facet() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        
        panel.background = element_rect(fill = 'white'),
        panel.grid = element_blank(),
        legend.position = 'none'
  )

ppi
```


```{r}
# setwd('C:/Users/patipark/Dropbox/STAR/miRNA')
# ggsave(ppi,
#        filename = 'Figure8_PPI.pdf',
#        height = 15,
#        width = 15,
#        units = 'cm',
#        device = cairo_pdf)
# dev.off()
```

Prepare input data.
```{r}
library(igraph)
library(plyr)
#transform df into igraph
vs <- graph.data.frame(my.edges, 
                      directed = FALSE, 
                      vertices = NULL)    

edgeCount <- lapply(1:length(my.edges$node1), function(i) {

  vert <- my.edges$node1[i]
  inc <- length(incident(vs, vert, mode = "total"))
  
  datout <- data.frame('Gene' = vert,
             'Edge.No' = inc)
  return(datout)

})

edgeCount <- do.call(rbind,edgeCount)

edgeCount <- edgeCount[!duplicated(edgeCount$Gene),]

edgeQ75 <- edgeCount %>% subset(Edge.No > 6)

T601.Q75 <- T601 %>% subset(Target.gene %in% edgeQ75$Gene)

T630.Q75 <- T630 %>% subset(Target.gene %in% edgeQ75$Gene)

```

```{r}
setwd('C:/Users/patipark/Dropbox/STAR/miRNA')
miRNA.PI <- read_tsv('miRNA_protein_interactions.tsv')
```

Plot miRNA-Protein interaction network.
```{r}
library(tidyverse)
## create edges dataframe
miR.edges <- miRNA.PI %>% 
  dplyr::select(node1,node2,combined_score)

# add miRNA to the edge table
miR.T601.F <- data.frame('node1' = 'hsa-miR-601',
                       'node2' = unique(miR.edges$node1),
                       'combined_score' = 0.9)

miR.T601.R <- data.frame('node1' = miR.T601.F$node2,
                    'node2' = miR.F$node1,
                    'combined_score' = 0.9)

miR.T630.F <- data.frame('node1' = 'hsa-miR-630',
                       'node2' = unique(miR.edges$node1),
                       'combined_score' = 0.9)

miR.T630.R <- data.frame('node1' = miR.T630.F$node2,
                    'node2' = miR.F$node1,
                    'combined_score' = 0.9)

miR.edges.cb <- rbind(miR.edges,miR.T601.F,miR.T601.R,miR.T630.F,miR.T630.R)

miR.vertices <- miR.edges.cb %>% dplyr::select(node1) %>% unique()

# part of codes is from 
# https://stackoverflow.com/questions/45003500/create-an-adjacency-matrix-or-list-based-on-shared-properties-from-a-data-frame
library(igraph)
# Create graph
g.miR <- graph_from_data_frame(d = miR.edges.cb, 
                           vertices = miR.vertices,
                           directed = TRUE)
# Create adjacency matrix from graph
M.miR <- as_adjacency_matrix(g.miR, attr = 'combined_score')

# coerces the data set as a matrix
m.miR <- as.matrix(M.miR) 
# this will create an 'igraph object'
gg.miR <- graph.adjacency(m.miR,
                      mode = 'undirected', 
                      weighted = 'combined_score' # use 'combined_socre' as weight
                      )

# change layout algorithm
# miR.layout <- layout.reingold.tilford(gg.miR, circular = TRUE)
# miR.layout <- layout.kamada.kawai(gg.miR)
miR.layout <- layout.circle(gg.miR, order = order(degree(gg.miR)))

# convert igraph object to coordinates for building a network using ggplot2
library(intergraph)
g2.miR <- asNetwork(gg.miR)
library(ggnetwork)
g3.miR <- ggnetwork(g2.miR,
                layout = miR.layout,
                cell.jitter = 0.75)

g3.1.miR <- g3.miR[!duplicated(g3.miR$vertex.names),]

# add node size according to degree of interaction
edge.degree <- degree(gg.miR)
g3.miR$degree <- edge.degree[match(g3.miR$vertex.names,names(edge.degree))]

# add node color
g3.miR$color <- ifelse(g3.miR$vertex.names %in% c('hsa-miR-601','hsa-miR-630'), 'orange','gray')

# order of node by edge number
# edgeQ75 <- edgeQ75 %>% arrange(desc(Edge.No))

# edge.order <- c('hsa-miR-601','hsa-miR-630',as.character(edgeQ75$Gene))

# add factor level 

```

Plot miRNA-Protein interaction network.
```{r}
library(ggplot2)
miR.ppi <- ggplot(g3.miR,
              aes(x = x, 
                  y = y, 
                  xend = xend, 
                  yend = yend)) +
  geom_edges(aes(linetype = 'solid'),
             #arrow = arrow(length = unit(6, "pt"), type = "closed") #if directed
             color = 'grey50',
             #curvature = 0.2,
             alpha = 0.5) +
  # geom_nodes(fill = 'gray', 
  #            shape = 21,stroke = 0.1,
  #            size = 3) + 
  geom_node_point(aes(size = degree,fill = color),
                  # fill = color,
                  shape = 21,
                  stroke = 0.1) + 
  scale_size(range = c(2,10)) +
  geom_text(data = g3.1.miR,
                  aes(x = x,
                      y = y,
                      label = vertex.names), 
                  size = 2.5,
                  segment.color = NA) +
  # scale_colour_manual(values = cols, name = 'Expresion level \n(z-score)') +
  # facet_grid(cellsub ~ cell, scales = 'fixed') + #, scales = 'free', space = 'free') +
  coord_fixed(ratio = 1) +
  # theme_facet() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        
        panel.background = element_rect(fill = 'white'),
        panel.grid = element_blank(),
        legend.position = 'none'
  )

miR.ppi

```

```{r}
# setwd('C:/Users/patipark/Dropbox/STAR/miRNA')
# ggsave(miR.ppi,
#        filename = 'Figure9_miR_PPI.pdf',
#        height = 12,
#        width = 12,
#        units = 'cm',
#        device = cairo_pdf)
# dev.off()
```
